### 2021-04-12 input/figadd.tex
preparing

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:addA}Both Process~$a$ and Process~$d$
      become sources.  $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]  {\input{input/figaddA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:addB}Messages transit through communication
      links and carry increasing weights.]{\input{input/figaddB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:addC}Process~$b$ and
      Process~$c$ receive, deliver, and forward $\alpha_{d}^{1}$ and
      $\alpha_d^3$ respectively.]{\input{input/figaddC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:addD}$a$ and
      $b$ discarded their received messages. $c$ still improved with $\alpha_d^2$.]
    {\input{input/figaddD.tex}}
    \caption{\label{fig:add}Efficient consistent partitioning using
      \NAMEB. Partition~$P_a$ includes $a$ while Partition~$P_d$
      includes $b$, $c$, and $d$. Process~$c$ and Process~$d$ never
      acknowledge the existence of Source~$a$, for Process~$b$ stops
      the propagation of the latter's notifications.}
  \end{center}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:addA}Both \Process~$a$ and \Process~$d$
      become sources.  $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]  {\input{input/figaddA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:addB}Messages transit through communication
      links and carry increasing weights.]{\input{input/figaddB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:addC}\Process~$b$ and
      \Process~$c$ receive, deliver, and forward $\alpha_{d}^{1}$ and
      $\alpha_d^3$ respectively.]{\input{input/figaddC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:addD}$a$ and
      $b$ discarded their received messages. $c$ still improved with $\alpha_d^2$.]
    {\input{input/figaddD.tex}}
    \caption{\label{fig:add}Efficient consistent partitioning using
      \NAMEB. Partition~$P_a$ includes $a$ while Partition~$P_d$
      includes $b$, $c$, and $d$. \Process~$c$ and \Process~$d$ never
      acknowledge the existence of Source~$a$, for \Process~$b$ stops
      the propagation of the latter's notifications.}
  \end{center}
\end{figure*}

### 2021-04-12 input/figdel.tex
rework issue with delete
meow
preparing

\begin{figure}
  \begin{center}
    \subfloat[Part A][\label{fig:delA}Process~$a$ deletes its
      partition.  It notifies all processes that belong to its
      partition using \NAMEB.]{\input{input/figdelA.tex}}
    \hspace{5pt}
    %
    \subfloat[Part B][\label{fig:delB}$\delta$s stop as soon as they
      encounter another partition. Process~$b$ answers with its
      partition.]{\input{input/figdelB.tex}}
  \end{center}
  \caption{\label{fig:del}Efficient removal of a partition using
    \NAME. Process~$a$ eventually knows that it belong to Partition
    $P_c$.}
\end{figure}

>>>>

\begin{figure}
  \begin{center}
    \subfloat[Part A][\label{fig:delA}\Process~$a$ deletes its
      partition.  It notifies all \processes that belong to its
      partition using \NAMEB.]{\input{input/figdelA.tex}}
    \hspace{5pt}
    %
    \subfloat[Part B][\label{fig:delB}$\delta$s stop as soon as they
      encounter another partition. \Process~$b$ answers with its
      partition.]{\input{input/figdelB.tex}}
  \end{center}
  \caption{\label{fig:del}Efficient removal of a partition using
    \NAMEB. \Process~$a$ eventually knows that it belong to Partition
    $P_d$.}
\end{figure}

### 2021-04-12 input/figproblem.tex
preparing

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both Process~$a$ and
      Process~$d$ add then delete their partition. For the recall,
      $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{5pt} \subfloat[Part B][\label{fig:problemB}Process~$b$
      delivers and forwards the $\alpha$ message about $P_a$ while
      Process~$c$ delivers and forwards both $\alpha$ and $\delta$
      messages about $P_d$.]  {\input{input/figproblemB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:problemC}Process~$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before. Process~$c$ and Process~$d$ do not
      deliver $\alpha_d$, for they already deleted
      it before.]{\input{input/figproblemC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:problemD}Processes
      discard remaining messages. Since Process~$b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted partition
      $P_a$.]{\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both \Process~$a$ and
      \Process~$d$ add then delete their partition. For the recall,
      $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{5pt} \subfloat[Part B][\label{fig:problemB}\Process~$b$
      delivers and forwards the $\alpha$ message about $P_a$ while
      \Process~$c$ delivers and forwards both $\alpha$ and $\delta$
      messages about $P_d$.]  {\input{input/figproblemB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:problemC}\Process~$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before. \Process~$c$ and \Process~$d$ do not
      deliver $\alpha_d$, for they already deleted
      it before.]{\input{input/figproblemC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:problemD}\Processes
      discard remaining messages. Since \Process~$b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted partition
      $P_a$.]{\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

### 2021-04-12 input/figproofB.tex
wip on implem mostly
preparing

  \newcommand\X{70pt}
  \newcommand\Y{-40pt}

@@

  \draw[opacity=0] (-1.5*\X, 0) -- (2.5*\X, 0);

  \draw[<->] (-5pt - \X, -\Y) -- node[above]{I} (5pt, -\Y); %% I as well
  \draw[->] (-\X + 5pt, 0) -- node[above, font=\small, color=\PC]{$\DEL_d$} (0 - 5pt, 0); %% d - e

  \draw[<->] (-5pt , 5pt -\Y) -- node[above]{III} (5pt + \X, 5pt -\Y); %% IV
  \draw[->] (0 +5pt, 0) -- (\X -5pt, 0); %% e - b'

  \draw[<->] (-5pt + \X , 10pt -\Y) -- node[above]{IV} (5pt + 2*\X, 10pt -\Y); %% V
  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% b - b''

@@

  \draw (-\X, 5pt) node[above, color=\PC]{$\DEL_d$};

  \draw ( 0, 5pt) node[above]{$\textcolor{\PC}{\ADD_d} \rightarrow \textcolor{\WRONG}{\ADD_f^{0.5}}$};
  \draw ( 0, -5pt) node[below, font=\footnotesize]{expect $\textcolor{\PB}{\DEL_f}$};

  \draw ( \X, 5pt) node[above]{$\textcolor{\PB}{\ADD_f^1} \rightarrow \textcolor{\PB}{\DEL_f} \rightarrow \textcolor{\PC}{\ADD_d^{0.5}}$};
  \draw ( \X, -5pt) node[below, font=\footnotesize]{expect $\textcolor{\PC}{\DEL_d}$ or $\textcolor{\PC}{\DEL_d^{0.5}}$};

  \draw (2*\X, 5pt) node[above]{$\textcolor{\PC}{\ADD_d^{1.5}}$};
  \draw (2*\X, -5pt) node[below, font=\footnotesize]{expect $\textcolor{\PC}{\DEL_d}$ or $\textcolor{\PC}{\DEL_d^{1.5}}$};

>>>>

  \newcommand\X{85pt}
  \newcommand\Y{-40pt}

@@

  \draw[opacity=0] (-1.4*\X, 0) -- (2.4*\X, 0);

  \draw[<->] (-5pt - \X, -\Y) -- node[above]{I': propagation} (5pt, -\Y); %% I as well
  \draw[->] (-\X + 5pt, 0) -- node[above, font=\small]{$\DEL_d$} (0 - 5pt, 0); %% d - e

  \draw[<->] (-5pt , 5pt -\Y) -- node[above]{III: detection} (5pt + \X, 5pt -\Y); %% IV
  \draw[->] (0 +5pt, 0) -- (\X -5pt, 0); %% e - b'

  \draw[<->] (-5pt + \X , 10pt -\Y) -- node[above]{IV: propagation} (5pt + 2*\X, 10pt -\Y); %% V
  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% b - b''

@@

  \draw (-\X, 5pt) node[above]{$\DEL_d$};

  \draw ( 0, 5pt) node[above]{$\textcolor{\PC}{\ADD_d} \rightarrow \textcolor{\WRONG}{\ADD_f^{0.5}}$};
  \draw ( 0, -5pt) node[below, font=\footnotesize]{expect $\DEL_f$};

  \draw ( \X, 5pt) node[above]{$\textcolor{\PB}{\ADD_f^1} \rightarrow \DEL_f \rightarrow \textcolor{\PC}{\ADD_d^{0.5}}$};
  \draw ( \X, -5pt) node[below, font=\footnotesize]{expect $\DEL_d$ or $\DEL_d^{0.5}$};

  \draw (2*\X, 5pt) node[above]{$\textcolor{\PC}{\ADD_d^{1.5}}$};
  \draw (2*\X, -5pt) node[below, font=\footnotesize]{expect $\DEL_d$ or $\DEL_d^{1.5}$};

### 2021-04-12 input/proposal.tex
preparing

\subsection{Dynamic consistent partitioning}

@@

Then, at any time, a source can revoke its self-appointed status of
source, hence deleting its partition from the system, by executing a
\texttt{Del} operation. All \processes that belong to this partition
must eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process $a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process $b$ -- that may belong
to its partition using \NAMEB. Upon receipt, \Process $b$ discards the
message $\delta_a$, for the latter does not target the former's
partition. Yet, \Process $b$ sends its own partition that may be the
best for \Process $a$. Eventually, every \processes belong to the same
partition $P_d$.

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
inconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time.

One could solve this issue by always propagating the $\delta$ messages
corresponding to the $\alpha$ messages it propagated before. In
Figure~\ref{fig:problem}, it means that as soon as \Process~$b$
forwards $\alpha_a^2$, it assumes that its neighbors \Process~$c$ and
\Process~$d$ may need the notification of removal $\delta_a$ if it
exists. However, such solution also implies that \processes generate
traffic not only related to their current partition, but also related
to partitions they belong to in the past. Instead, we propose to use
the delivery order at each \process to detect possible inconsistencies
and solve them. Together, \processes eventually remove all stale
control information (transiting messages and local states) of the
system leaving room for propagation of messages about up-to-date
partitions.

@@

\begin{corollary}[\label{theo:dcp}Dynamic consistent partitioning]
  Consistent partitioning where each \process can \texttt{Add} and
  \texttt{Del} its partition at any time requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, then
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}

\begin{proof}
  \TODO{Quite obvious.}
\end{proof}

We define a last win order to forbid the delivery of stale information
emanating from a same \process. In other terms, the \emph{best}
notification message is always the most recent known of each \process.

\begin{definition}[Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$.
\end{definition}

@@

\begin{theorem}[\TODO{meow}]
  \TODO{Puring stale notifications requires to handle I...IV.}
\end{theorem}

@@

@@

We extend Algorithm~\ref{algo:add} to enable each source to revoke its
self-appointed status of source and become a simple \process again.
Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

>>>>

\subsection{Dynamic consistent partitioning}
\label{subsec:dynamic}

@@

At any time, a source can revoke its self-appointed status of source,
hence deleting its partition from the system, by executing a
\texttt{Del} operation. All \processes that belong to this partition
must eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process $a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process $b$ -- that may
belong to its partition using \NAMEB. Upon receipt, \Process $b$
discards the message $\delta_a$, for the latter does not target the
former's partition. \Process $b$ sends its own knowledge of best
partition $\alpha_d^3$ that may prove to be the best for \Process
$a$. Eventually, every \processes belong to the same partition
$P_d$. In this scenario, they reach consistent partitioning.

\texttt{Del}ete operations add a new notion of order between events,
and most importantly between message deliveries. Delete operations
implicitly state that all preceding events become obsolete, and that
all messages originating from these preceding events convey stale
control information.

\begin{definition}[Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m'$
  convey \emph{stale} control information.
\end{definition}

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
inconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time. \Processes need to exploit their local knowledge gathered over
message receipts.

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. Instead, we propose to use the delivery order at each \process
to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\begin{corollary}[\label{theo:dcp}Dynamic consistent partitioning]
  Consistent partitioning where each \process can \texttt{Add} and
  \texttt{Del}ete its partition at any time requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. The only case where a \process would block such epidemic
  propagation would be when it received and delivered a message from a
  partition that has since then been deleted but with a better sum of
  weights than newly received ones. Removing \emph{forever} all such
  stale messages about deleted partitions would allow \processes to
  propagate their best up-to-date partition again, eventually reaching
  consistent partition together.
\end{proof}

@@

\begin{theorem}[Propagation $+$ detection $\implies$ DCP]
  % Propagating 
  \TODO{Puring stale notifications requires to handle I...IV.}
\end{theorem}

@@

Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

@@
